### 2019-06-19 2-Chapters/3-Chapter/chapter3.tex
Plots for end of Chapter 3

\graphicspath{{2-Chapters/3-Chapter/Images/}}
\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}

@@

\textbf{References.}
%
The code for SMPyBandits is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}}, and all the library is freely publicly released under the MIT open-source License.
This chapter is based on our article \cite{SMPyBanditsJMLR}.

@@

@@

We show how to install SMPyBandits, and an example of how to run a simple experiment.
This bash code\footnote{~See this page \texttt{\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}} for more details.} shows how to clone the code\footnote{~SMPyBandits is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}. You can install it directly with \texttt{sudo pip install SMPyBandits}, or from a \texttt{virtualenv} \cite{virtualenv}.},
and install the requirements for Python 3 (once):

@@

% https://tex.stackexchange.com/a/12430/
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
$ for T in $(seq 5000 5000 50000); do \
    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
        N=1000  BAYES=True    T=$T         K=8 \
        python3 main.py configuration.py \
    && cp logs/main_py3_log.txt \
          logs/main_py3_log__N1000_BAYES_T{T}_K8.txt \
  done
$ for K in $(seq 2 2 32); do \
    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
        N=1000  BAYES=True    T=5000       K=$K \
        python3 main.py configuration.py \
    && cp logs/main_py3_log.txt \
          logs/main_py3_log__N1000_BAYES_T5000_K{K}.txt \
done
    \end{minted}
    \caption{Bash code to run the large-scale experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}, for $K=8$ and $T\in\{5000,\dots,50000\}$ and $T=5000$ and $K\in\{2,\dots,32\}$.}
    % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:BashCodeToLaunchLargeExperiments}.}
    \label{lst:3:BashCodeToLaunchLargeExperiments}
\end{listing}

@@

\TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ? Ce n'est pas crucial.
    % C'est en ligne sur \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
}

@@

% https://tex.stackexchange.com/a/12430/
\begin{small}
    \begin{listing}[h!]
        \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=1 \
  python3 main.py configuration.py  # on only one core!

$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=-1 \
  python3 main.py configuration.py  # parallel on all the cores!

$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=100 \
  python3 main.py configuration.py  # parallel with too many jobs!
        \end{minted}
        \caption{Running a simulation using either one CPU core with \texttt{N\_JOBS=1}, or all of them with \texttt{N\_JOBS=-1}, or more jobs than CPU cores with \texttt{N\_JOBS=20} (if there are $4$ cores). The later option is not advised, and \texttt{N\_JOBS=-1} gives the optimal speed-up factor.}
        % \captionof{listing}{Run a simulation using 1 or \texttt{N} cores \label{lst:3:runOneCoreOrMore}.}
        \label{lst:3:runOneCoreOrMore}
    \end{listing}
\end{small}

\textbf{Effect of the number of jobs.}
%
As long as the number of jobs (\texttt{N\_JOBS} here) is less then or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal.
%  dividing the running time by \texttt{N\_JOBS}.
But jobs are implemented as threads, so the speed-up cannot be more than the number of cores, and using for instance $20$ jobs on $4$-cores for the $20$ repetitions is sub-optimal, as the CPU will essentially spend all its time (and memory) managing the different jobs, and not actually doing the simulations.
Using the above example, we illustrate the effect of using multi-jobs and multi-cores on the time efficiency of simulations using SMPyBandits. We consider three values of \texttt{N\_JOBS}, $1$ to use only one core and one job, $4$ to use all the $4$ cores of my $i5$ Intel CPU, and $20$ jobs.
%
We give in Table~\ref{table:3:speedUpTimeParallelComputations} an example of running time of an experiment with $T=10000$, and different number of repetitions and number of jobs.
These three experiments correspond to the three lines of Code Example~\ref{lst:3:runOneCoreOrMore} shown above.
%
The results of Table~\ref{table:3:speedUpTimeParallelComputations} and of Figure~\ref{fig:3:analyze_speedup_as_nb_of_jobs}
clearly illustrate that using more jobs than the number of CPU is sub-optimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
Due to the cost of orchestrating the different jobs, and memory exchanges at the end of each repetition, the parallel version is \emph{not} $4$ times faster, but empirically we always found it to be $2$ to $3.5$ times faster.
%
More details are included in the documentation page
\href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}

@@

\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|rrr}
    \textbf{Repetitions} & \texttt{N\_JOBS}$=1$ & \textcolor{darkgreen}{\texttt{N\_JOBS}$=4$ ($= n_{\text{cores}}$)} & \texttt{N\_JOBS}=$20$ ($> n_{\text{cores}}$) \\
        \hline
        $N=1$ repetition    & \SI{15}{\second}
        & \SI{26}{\second} ($0.57 \%$)  % 15/26
        & \SI{43}{\second} ($0.34 \%$)  % 15/43
        \\
        $N=10$ repetitions  & \SI{87}{\second}
        & \SI{51}{\second} ($1.70 \%$)  % 87/51
        & \SI{76}{\second} ($1.14 \%$)  % 87/76
        \\
        $N=100$ repetitions & \SI{749}{\second}
        & \SI{272}{\second} ($2.75 \%$)  % 749/272
        & \SI{308}{\second} ($2.43 \%$)  % 749/308
        \\
        $N=500$ repetitions & \SI{2944}{\second}
        & \SI{1530}{\second} ($1.92 \%$)  % 2944/1530
        & \SI{1846}{\second} ($1.59 \%$)  % 2944/1846
        \\
        % \hline
    \end{tabular}
    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, on a machine with $n_{\text{cores}}=12$ CPU cores, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$. The numbers in parentheses are the absolute gain in comparison with the non-parallel simulation (first column), in percentages.}
    \label{table:3:speedUpTimeParallelComputations}
% \end{footnotesize}
\end{table}

\begin{figure}[h!]  % [htbp]
	\centering
	% \includegraphics[width=0.75\linewidth]{logs/analyze_speedup_as_nb_of_jobs.pdf}
	\includegraphics[width=0.75\linewidth]{logs/analyze_speedup_as_nb_of_jobs_normalize.pdf}
	\caption{Absolute gain of the running time, for the simulation of Code Example~\ref{lst:3:runOneCoreOrMore}, as a function of the number of jobs \texttt{N\_JOBS}, for a $n_{\text{cores}}=12$-core machine.}
	\label{fig:3:analyze_speedup_as_nb_of_jobs}
\end{figure}

>>>>

\graphicspath{{2-Chapters/3-Chapter/Images/}{2-Chapters/3-Chapter/logs/}{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}

@@

\paragraph{References.}
%
The code for SMPyBandits is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}}, and all the library is freely publicly released under the MIT open-source License.
This chapter is based on our article \cite{SMPyBanditsJMLR}.

@@

% FIXME?
\vfill{}
% FIXME?

@@

We show how to install SMPyBandits, and an example of how to run a simple experiment.
See this page \texttt{\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}} for more details.
SMPyBandits is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}, you can install it directly with \texttt{sudo pip install SMPyBandits}, or from a \texttt{virtualenv} \cite{virtualenv}.
This bash code shows how to clone the code, and install the requirements for Python 3 (once):

@@

% https://tex.stackexchange.com/a/12430/
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
$ for T in $(seq 5000 5000 50000); do \
    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 N=1000 BAYES=True \
    T=$T   K=8  python3 main.py configuration.py \
    && cp logs/main_py3_log.txt logs/main_py3_log__N1000_BAYES_T{T}_K8.txt \
  done
$ for K in $(seq 2 2 32); do \
    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 N=1000 BAYES=True \
    T=5000 K=$K python3 main.py configuration.py \
    && cp logs/main_py3_log.txt logs/main_py3_log__N1000_BAYES_T5000_K{K}.txt \
  done
    \end{minted}
    \caption{Bash code to run the large-scale experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}, for $K=8$ and $T\in\{5000,\dots,50000\}$ and $T=5000$ and $K\in\{2,\dots,32\}$.}
    % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:BashCodeToLaunchLargeExperiments}.}
    \label{lst:3:BashCodeToLaunchLargeExperiments}
\end{listing}

@@

% \TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ? Ce n'est pas crucial.
%     % C'est en ligne sur \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
% }

@@

% https://tex.stackexchange.com/a/12430/
\begin{small}
    \begin{listing}[h!]
        \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=1 \
  python3 main.py configuration.py  # on only one core

$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=-1 \
  python3 main.py configuration.py  # parallel on all the cores

$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=100 \
  python3 main.py configuration.py  # parallel with too many jobs
        \end{minted}
        \caption{Running a simulation using either one CPU core with \texttt{N\_JOBS=1}, or all of them with \texttt{N\_JOBS=-1}, or more jobs than CPU cores with \texttt{N\_JOBS=20} (if there are $4$ cores). The later option is not advised, and \texttt{N\_JOBS=-1} gives the optimal speed-up factor.}
        % \captionof{listing}{Run a simulation using 1 or \texttt{N} cores \label{lst:3:runOneCoreOrMore}.}
        \label{lst:3:runOneCoreOrMore}
    \end{listing}
\end{small}

\textbf{Effect of the number of jobs.}
%
As long as the number of jobs (\texttt{N\_JOBS} here) is less then or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal.
%  dividing the running time by \texttt{N\_JOBS}.
But jobs are implemented as threads, so the speed-up cannot be more than the number of cores, and using for instance $20$ jobs on $4$-cores for the $20$ repetitions is sub-optimal, as the CPU will essentially spend all its time (and memory) managing the different jobs, and not actually doing the simulations.
Using the above example, we illustrate the effect of using multi-jobs and multi-cores on the time efficiency of simulations using SMPyBandits. We consider three values of \texttt{N\_JOBS}, $1$ to use only one core and one job, $4$ to use all the $4$ cores of my $i5$ Intel CPU, and $100$ jobs.
%
We give in Table~\ref{table:3:speedUpTimeParallelComputations} an example of running time of an experiment with $T=10000$, and different number of repetitions and number of jobs.
These three experiments correspond to the three lines of Code Example~\ref{lst:3:runOneCoreOrMore} shown above.
%
The results of Table~\ref{table:3:speedUpTimeParallelComputations} clearly illustrate that using more jobs than the number of CPU is sub-optimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
Due to the cost of orchestrating the different jobs, and memory exchanges at the end of each repetition, the parallel version is \emph{not} $4$ times faster, but empirically we always found it to be $2$ to $3.5$ times faster.
%
Similar conclusions can be drawn from Figure~\ref{fig:3:analyze_speedup_as_nb_of_jobs}, which reports the results of the same experiment on a machine with $n_{\text{cores}}=12$ CPU cores.
More details are included in the documentation page
\href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}

@@

\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|rrr}
    \textbf{Repetitions} & \texttt{N\_JOBS}$=1$ & \textcolor{darkgreen}{\texttt{N\_JOBS}$=4$ ($= n_{\text{cores}}$)} & \texttt{N\_JOBS}=$20$ ($> n_{\text{cores}}$) \\
        \hline
        $N=1$ repetition    & \SI{15}{\second}
        & \SI{26}{\second} ($0.57 \%$)  % 15/26
        & \SI{43}{\second} ($0.34 \%$)  % 15/43
        \\
        $N=10$ repetitions  & \SI{87}{\second}
        & \SI{51}{\second} ($1.70 \%$)  % 87/51
        & \SI{76}{\second} ($1.14 \%$)  % 87/76
        \\
        $N=100$ repetitions & \SI{749}{\second}
        & \SI{272}{\second} (\textcolor{darkgreen}{$2.75 \%$})  % 749/272
        & \SI{308}{\second} ($2.43 \%$)  % 749/308
        \\
        $N=500$ repetitions & \SI{2944}{\second}
        & \SI{1530}{\second} (\textcolor{darkgreen}{$1.92 \%$})  % 2944/1530
        & \SI{1846}{\second} ($1.59 \%$)  % 2944/1846
        \\
        % \hline
    \end{tabular}
    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, on a machine with $n_{\text{cores}}=4$ CPU cores, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$. The percentages are the relative gains in comparison with the non-parallel simulation.}
    \label{table:3:speedUpTimeParallelComputations}
% \end{footnotesize}
\end{table}

\begin{figure}[h!]  % [htbp]
	\centering
	% \includegraphics[width=0.85\linewidth]{analyze_speedup_as_nb_of_jobs.pdf}
	\includegraphics[width=0.95\linewidth]{analyze_speedup_as_nb_of_jobs_normalize.pdf}
	\caption{Relative gain on the running time in comparison to using one job, as a function of the number of jobs \texttt{N\_JOBS}, for a $n_{\text{cores}}=12$-core machine running the simulation of Code Example~\ref{lst:3:runOneCoreOrMore}.}
	\label{fig:3:analyze_speedup_as_nb_of_jobs}
\end{figure}
